---
title: "GOM Contaminated Sediments Analysis: Data Tidying"
author: "Joshua Harkness and Autumn Pauly,"
date: "2023-10-20"
output: github_document
---
# Introduction
This document is tidying the data sets of a series of reports published by the U.S. Geological Survey about the distributions and concentrations of contaminated sediments in the Gulf of Maine. 

## Load packages
These are the packages that will be needed to tidy data frames.
```{r load-packages}
library(tidyverse)
library(readxl)
library(sf)
library(visdat)
library(naniar)
```

##Loading Full Data Sets
These data sets are very large. Since these are all very large data files, we will select useful variables and variables we are interested in looking at to pare them down to a more manageable size.

```{r load-data, warning=FALSE}
stations_full = read_xls("/cloud/project/data/original_datasets/STAT2002.xls", sheet = 2, skip = 3)
PCBs_full = read_excel("/cloud/project/data/original_datasets/PCBP2002.xls", sheet = 2, skip = 3)
PAHs_full = read_excel("/cloud/project/data/original_datasets/PAHS2002.xls", sheet = 2, skip = 3)
organics_full = read_excel("/cloud/project/data/original_datasets/GENO2002.xls", sheet = 2, skip = 3)
```

#Selecting and Modifying Variables 
###PCBs
Creates PCBs2002 data frame. These are variables of interest, principally contaminant concentration with assigned unique IDs that can be associated to unique IDs in the other data frames.

```{r select-var-PCBs}
PCBs2002 = select(PCBs_full, c(UNIQUE_ID, PCB_52_NGG, PCB101_NGG, PCB118_NGG, PCB128_NGG, PCB138_NGG, PCB153_NGG, PCB180_NGG, PCB206_NGG, PCB209_NGG, DDT_4_4_C, DDT_2_4_C, DDE_4_4_C, DDD_4_4_C, ENDRIN_C, ENDR_ALD_C, ALDRIN_C, DIELDRN_C, CLRDNE_T_C, MIREX_C, METHOXYCLC, BHC_A_C, BHC_B_C, BHC_D_C, LINDANE_C))

```

###Stations
Creates Stations2002 data frame.  These are unique IDs, coordinates (Lat/Lon in decimal degrees), depth sounding (meters), state name, USGS quad name, general and specific location names, area code, sampling date, resampling date, and information on sediment collection. 

```{r select-var-STATs}
Station2002 = select(stations_full, c(UNIQUE_ID, LATITUDE, LONGITUDE, SOUNDING_M, STATE_NAME, QUAD_NAME, GEN_LOC_NM, SPECFC_LOC, AREA_CODE, SAMP_DATE1, TO_SMP_DT2, DPTH_N_COR, DPTH_CODE, COR_GRB_CD))

Station2002 <- Station2002 %>%
mutate(GEN_LOC_NM = fct_recode(GEN_LOC_NM,
                           "Rockland to north" = "North of 44; to 50M isobath",
                           "Cape Elizabeth to Rockland" = "43.5N to 44N; to 50M isobath", 
                           "Cape Ann to Cape Elizabeth" = "CAPE ANN to 43.5N", 
                           "Boston Inner Harbor" = "BOSTON INNER HARBOR", 
                           "Cape Code Bay" = "CAPE COD BAY",
                           "Gulf of Maine (>50m Isobath)" = "GULF OF MAINE, >50M ISOBATH",
                           "Northwest Boston Harbor" = "NORTHWEST BOSTON HARBOR",
                           "Southeast Boston Harbor" = "SOUTHEAST BOSTON HARBOR",
                           "Gulf of Maine (<=50m Isobath)" = "GULF OF MAINE, <=50M",
                           "Central Boston Harbor" = "CENTRAL BOSTON HARBOR",
                           "Inland/Rivers" = "INLAND / RIVERS",
                           "Harbor Approaches" = "HARBOR APPROACHES",
                           "Massachusetts Bays" = "MASS BAYS"))

Station2002 <- Station2002 %>%
  mutate(site = fct_recode(SPECFC_LOC,
                           "Chelsea River" = "Upper Chelsea River",
                           "Chelsea River" = "Lower Chelsea River", 
                           "Chelsea River" = "Chelsea River mouth",
                           "Chelsea River" = "Chelsea River A",
                           "Chelsea River" = "Chelsea River B",
                           "Chelsea River" = "Chelsea River C",
                           "Chelsea River" = "Chelsea River D",
                           "Chelsea River" = "Chelsea River E",
                           "Chelsea River" = "Chelsea River F",
                           "Chelsea River" = "Chelsea River G",
                           "Chelsea River" = "Chelsea River H",
                           "Chelsea River" = "Chelsea River, Gulf Oil Fuel Off-Loading Pier",
                           "Chelsea River" = "Chelsea River, Golf Oil Fuel Off-Loading Pier",
                           "Mystic River" = "MYSTIC RIVER",
                           "Mystic River" = 'Mystic River "B"',
                           "Mystic River" = 'Mystic "A"-7830',
                           "Mystic River" = 'Mystic "B"-7831',
                           "Chelsea River" = 'Chelsea "A"-7828',
                           "Chelsea River" = 'Chelsea "E"-7829',
                           "Island End River" = "ISLAND END RIVER",
                           "Weymouth Fore River" = "FORE RIVER", 
                           "Fore River" = "Portland Fore River", 
                           "Fore River" = "Portland Harbor & Fore River", 
                           "Saco River" = "Saco River & Camp Ellis Harbor", 
                           "Kennebec River" = "Kennebec River to Bath", 
                           "Kennebec River" =   "KENNEBEC RIVER",
                           "Kennebec River" = "KENNEBEC RIVER,ME.", 
                           "Penobscot River" = "PENOBSCOT RIVER", 
                           "Penobscot River" = "MOUTH PENOBSCOT RIVER",
                           "Penobscot River" = "Penobscot River To Bangor", 
                           "Fore River" = "Entrance to Weymouth Fore River near Buoy N-22", 
                           "Fore River" = "Entrance to Weymouth Fore River near Buoy N-28", 
                           "Fore River" = "Entrance to Weymouth Fore River near C-19", 
                            "Fore River" = "Weymouth Fore & Town River",
                            "Fore River" = "Weymouth Fore River",
                            "Fore River" = "Weymouth Fore River - Citgo's Braintree terminal",
                            "Fore River" = "Weymouth Fore Rvr",
                            "Fore River" = "WFR",
                           "Piscataqua River" = "In Piscataqua River NE of Outer Cutts Cove",
                           "Piscataqua River" = "E of Freeman's Pt. near mouth of inlet",
                           "Piscataqua River" = "Back Channel just south of Wattlebury Is.",
                           "Piscataqua River" = "Off dry dock 3 at PNSY 60' from pier",
                           "Piscataqua River" = "SE of Dry Dock 1 at PNSY 75' from pier",
                           "Piscataqua River" = "S. of Dry Dock 2 at PNSY",
                           "Piscataqua River" = "Piscataqua River N of Pierce Is.",
                           "Piscataqua River" = "Little Harbor S. of Shapleigh Island",
                           "Piscataqua River" = "SSE of pier of US Coast Guard station",
                           "Piscataqua River" = "Clark Island Embayment between Jamaica Is. and Hick Rocks",
                           "Piscataqua River" = "Clark Island Embayment S. of Jamaica Is.",
                           "Piscataqua River" = "Clark Island Embayment, Clark cove near Norway ball #17",
                           "Piscataqua River" = "Clark Island Embayment, W of Clark Is. pier",
                           "Piscataqua River" = "Clark Island Embayment N. of Clark Is. Causeway",
                           "Piscataqua River" = "S of Clark Island causeway",
                           "Piscataqua River" = "Back Channel E side of back gate entr. to PNSY",
                           "Piscataqua River" = "West of Jamaica Is. in Back Channel of PNSY",
                           "Piscataqua River" = "E. of Gooseberry Island",
                           "Spruce Creek" = "Spruce Creek E. of Admiralty Village",
                           "Spruce Creek" = "Back Channel Near N. shore of entrance to Barters Creek"
                           ))
```
"Piscataqua River" = ""
###Organics
Creates Organics2002 dataframe with selected variables of interest.  These include general organics as well as total PCB and some notable pesticide concentrations.

```{r select-var-ORG}
Organics2002 = select(organics_full, c(UNIQUE_ID, O_G_UGG, PHCTOT_UGG, PCB_T_UGG, DDT_T_NGG, DDE_T_NGG, DDD_T_NGG, PEST_UG_G, LIPIDS_NGG, CLOST_SP_G, MBT_C, DBT_C, TBT_C, TTBT_C))
```

###PAHs
Creates PAHs2002 dataframe with selected variables of interest.  These are principally contaminate concentrations with unique IDs associated with other data frames.

```{r select-var-PAH}
PAHs2002 = select(PAHs_full, c("UNIQUE ID", BENZNE_C, "2BZTPN_C", "2BZFRN_T_C", BIPHENYL_C, ACNPHTHN_C, ACNPHTYL_C, FLUORENE_C, PHNANTHR_C, ANTHRACN_C,  PYRENE_C, FLORNTHN_C, CHRYSENE_C, PERYLENE_C)) 

PAHs2002 <- PAHs2002 %>%
  rename("UNIQUE_ID" = "UNIQUE ID", 
         "BENZNE" = "BENZNE_C", 
         "2BZTPN" = "2BZTPN_C", 
         "2BZFRN" = "2BZFRN_T_C", 
         "BIPHENYL" = "BIPHENYL_C", 
         "ACNPHTHN" = "ACNPHTHN_C", 
         "ACNPHTYL" = "ACNPHTYL_C", 
         "FLUORENE" = "FLUORENE_C", 
         "PHNANTHR" = "PHNANTHR_C", 
         "ANTHRACN" = "ANTHRACN_C",  
         "PYRENE" = "PYRENE_C", 
         "FLORNTHN" = "FLORNTHN_C", 
         "CHRYSENE" = "CHRYSENE_C", 
         "PERYLENE" = "PERYLENE_C")

```

## Modifying Dataframes
### Joining Stations with Contaminent Data Sets
This action associates location information and contaminant data by each sample's `UNIQUE_ID`. 

```{r join-data-frames}
  PCBs_loc <- full_join(Station2002, PCBs2002, by = "UNIQUE_ID")
  
  PAHs_loc <- full_join(Station2002, PAHs2002, by = "UNIQUE_ID")
  
  Organics_loc <- full_join(Station2002, Organics2002, by = "UNIQUE_ID")
```

###Recoding PAHs' Special Character Values as Numeric
The amount values for the PAH containment `CHRYSENE` needs to be recoded from character values to numeric values. 

``` {r converting-PAHs-as.numeric}
  PAHs_loc <- PAHs_loc %>%
    mutate(CHRYSENE = recode(CHRYSENE, ">" = "NA")) 

  PAHs_loc %>%
    transform(PAHs_loc, CHRYSENE = as.numeric(CHRYSENE))
```

# Visualizing Missing Data
Visualizing missing data can help us determine which specific contaminants were actually detected, and should then be analyzed, as there are a lot of missing values within each data set.

###Missing PCB Data
``` {r missing-data-PCBs}
#Visualizing the percentage of missing data
#visdat::vis_miss(PCBs_loc)

#visualizing correlations in data as a heatmap - only numeric values...
#dplyr::select(PCBs_loc, where(is.numeric)) %>%
#visdat::vis_cor(PCBs_loc, "pearson", "pairwise.complete.obs", facet)

#we should highlight the locations where there are PCBs detected
```

```{r}
 visdat::vis_dat(PCBs_loc)
```

```{r}
naniar::gg_miss_var(PCBs_loc)
```

```{r}
naniar::miss_var_summary(PCBs_loc)
```


###Missing PAH Data
``` {r missing-data-PAHs}
#Visualizing the percentage of missing data
#visdat::vis_miss(PAHs_loc)

#visualizing correlations in data as a heatmap - only numeric values...
#dplyr::select(PAHs_loc, where(is.numeric)) %>%
#visdat::vis_cor(PAHs_loc, "pearson", "pairwise.complete.obs", facet)

#we should highlight the locations where there are PAHs detected
```

###Missing Organics Data
``` {r missing-data-Organics}
#Visualizing the percentage of missing data
# visdat::vis_miss(Organics_loc)

#visualizing correlations in data as a heatmap - only numeric values...
#dplyr::select(Organics_loc, where(is.numeric)) %>%
#visdat::vis_cor(Organics_loc, "pearson", "pairwise.complete.obs", facet)

#This shows us in which station specific organics are found
```

# Write loc datasets as CSVs
``` {r saving-data_loc-into-csv-files}
write_csv(Organics_loc, "/cloud/project/data/datasets_csv/Organics_loc.csv")
write_csv(PAHs_loc, "/cloud/project/data/datasets_csv/PAHs_loc.csv")
write_csv(PCBs_loc, "/cloud/project/data/datasets_csv/PCBs_loc.csv")
```