---
title: 'GOM Contaminated Sediments analysis: PAHs'
author: "Joshua Harkness and Autumn Pauly"
date: "2023-10-28"
output: github_document
---

```{r load-packages}
library(tidyverse)
library(sf)
library(leaflet)
library(ggplot2)
library(RColorBrewer)

```

#Loading data
```{r load-data}
PAHs <- read.csv(paste0("/cloud/project/data/datasets_loc/PAHs_loc.csv"), header = T)
```

#Glimpsing PAH
```{r glimpse-data, warning=FALSE}
glimpse(PAHs)
```

# Joining the Station Information to the PAH dataset
```{r joining-station-info, warning=FALSE}

PAHs_station <- full_join(Station2002, PAHs, by = "UNIQUE_ID")

PAHs_station <- PAHs_station %>%
  mutate(CHRYSENE_C = as.numeric(CHRYSENE_C)) %>%
  pivot_longer(cols = `BENZNE_C`:`B_GHI_PYLC`, 
               names_to = "pah_detected", 
               values_to = "amount_detected")
  
PAHs_station_nona <- PAHs_station %>%
   mutate(amount_detected_with_na = ifelse(is.na(amount_detected), 0, amount_detected))

PAHs_station = select(PAHs_station, c(UNIQUE_ID, LATITUDE.x, LONGITUDE.x, SOUNDING_M.x, QUAD_NAME.x, GEN_LOC_NM.x, SPECFC_LOC.x, AREA_CODE.x, SAMP_DATE1.x, TO_SMP_DT2.x, DPTH_N_COR.x, DPTH_CODE.x, COR_GRB_CD.x, pah_detected, amount_detected))

glimpse(PAHs_station)

```
Joining the station data with the PAH sample taken. The dataset is pivoted so that each observation is the PAH that was observed. Creates the PAHs_station_nona data frame that has the stations joined to the dataset with all na values omitted from the amount_detected column. 

# Plotting PAHs Detected in GOM
``` {r plotting-pahs-detected, warning=FALSE}

PAHs_station_nona %>%
  filter(pah_detected %in% c("BENZNE_C", "BENZNE_C", "NAPHTHLN_C", "BIPHENYL_C", "ACNPHTHN_C", "ACNPHTYL_C", "FLUORENE_C", "PHNANTHR_C", "ANTHRACN_C", "PYRENE_C", "FLORNTHN_C", "CHRYSENE_C", "PERYLENE_C")) %>%
  ggplot(aes(x = pah_detected, y = amount_detected, color = pah_detected)) +
  geom_col() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#I do now know why it is colored like that

```

``` {r plotting-names-test}

PAHs_station %>%
  ggplot(aes(x = LONGITUDE.x, y = LATITUDE.x, color = GEN_LOC_NM.x)) + 
  geom_point()

```

``` {r summary-stats-PAHs}
Summary_PAHs <- PAHs_station %>%
  group_by(GEN_LOC_NM.x) %>%
  drop_na(amount_detected) %>%
  summarise(mean_PAH = mean(amount_detected),
    sd_PAH = sd(amount_detected),
    n_PAH = n(),
    SE_PAH = sd(amount_detected) / sqrt(n()))

Summary_PAHs

```

``` {r pah-gen-loc}
Summary_PAHs %>%
  ggplot(aes(x = fct_rev(fct_reorder(GEN_LOC_NM.x, mean_PAH)), y = mean_PAH, fill = GEN_LOC_NM.x))+
  geom_col(col = "black")+
  geom_errorbar(aes(ymin = mean_PAH - SE_PAH, ymax = mean_PAH + SE_PAH), width = 0.2)+
  coord_flip()+
  scale_fill_viridis_d()+
  theme_bw()+
  theme(legend.position = "none")+
  labs(title = "PAH Concentrations in Gulf of Maine Sediments",
       x = "General location",
       y = "Mean total PAH concentration ug/g",
       caption = "Error bars = 1 standard error")
```
```{r pah-boston-harbor1}
Summary_PAHs %>%
  filter(GEN_LOC_NM.x %in% c("BOSTON INNER HARBOR", "CENTRAL BOSTON HARBOR", "SOUTHEAST BOSTON HARBOR", "NORTHWEST BOSTON HARBOR")) %>%
  ggplot(aes(x = fct_rev(fct_reorder(GEN_LOC_NM.x, mean_PAH)), y = mean_PAH, fill = GEN_LOC_NM.x))+
  geom_col(col = "black")+
  geom_errorbar(aes(ymin = mean_PAH - SE_PAH, ymax = mean_PAH + SE_PAH), width = 0.2)+
  scale_fill_brewer(type = "qual", palette = 4, direction = 1, aesthetics = "fill")+
  theme_bw()+
  coord_flip()+
  theme(legend.position = "none")+
  labs(title = "PAH Concentrations in Boston Harbor Sediments",
       x = "General location",
       y = "Mean total PAH concentration ug/g",
       caption = "Error bars = 1 standard error")
```
```{r pah-boston-harbor2}
Summary_PAHs %>%
  filter(GEN_LOC_NM.x %in% c("CENTRAL BOSTON HARBOR", "NORTHWEST BOSTON HARBOR")) %>%
  ggplot(aes(x = fct_reorder(GEN_LOC_NM.x, mean_PAH), y = mean_PAH, fill = GEN_LOC_NM.x))+
  geom_col(col = "black")+
  geom_errorbar(aes(ymin = mean_PAH - SE_PAH, ymax = mean_PAH + SE_PAH), width = 0.2)+
  scale_fill_brewer(type = "qual", palette = 4, direction = 1, aesthetics = "fill")+
  theme_bw()+
  theme(legend.position = "none")+
  labs(title = "PAH Concentrations in Boston Harbor Sediments",
       subtitle = "Central and Northwest Boston Harbor",
       x = "General location",
       y = "Mean total PCB concentration ug/g",
       caption = "Error bars = 1 standard error")
```

### Specific Locations


```{r}
Sum_PAH_site <- PAHs_station %>%
  group_by(SPECFC_LOC.x) %>%
  drop_na(amount_detected) %>%
  summarise(mean_PAH = mean(amount_detected),
    sd_PAH = sd(amount_detected),
    n_PAH = n(),
    SE_PAH = sd(amount_detected) / sqrt(n()))

Sum_PAH_site
```
