---
title: 'GOM Contaminated Sediments Analysis: PAHs'
author: "Joshua Harkness and Autumn Pauly"
date: "2023-10-28"
output: github_document
---
# Introduction
This document is an analysis of distributions and concentrations of Polycyclic aromatic hydrocarbons (PAH) contaminants that were reported in a publication by the U.S. Geological Survey on contaminated sediments in the Gulf of Maine. In this data set, we analyzed the prevalence of Benzene, Dibenzothiophene, Dibenzofuran, Biphenyl, Acenaphthene, Acenaphthylene, Fluorene, Phenanthrene, Anthracene,  Pyrene , Fluoranthene, Chrysene, and Perylene. 

# Background
Polycyclic aromatic hydrocarbons (PAH) are a class of organic compounds that are composed of multiple aromatic rings (also known as cyclic rings) made up of carbon and hydrogen atoms. These contaminants are formed during the incomplete combustion of organic materials such as fossil fuels, wood, and other organic matter. They can be found in various environmental matrices, including air, soil, water, and sediments. Common sources of PAHs include vehicle exhaust, industrial processes, burning of coal, oil, wood, and tobacco smoke.

The Agency for Toxic Substances and Disease Registry states that PAHs have a low degree of acute toxicity to humans. When chronically exposed to high doses, PAH toxicity can induce issues within the pulmonary, gastrointestinal, renal, and dental systems as well as increase an individual's risk of cancer (Mumtaz et al. 1996). The International Agency for Research on Cancer (IARC) has classified certain PAHs, such as benzo[a]pyrene, as Group 1 human carcinogens (IARC, 2010). In regards to the environment, PAHs have moderate to high acute toxicity to wildlife, specifically aquatic life and birds. Research has found that PAHs have toxic compounds that affect immunotoxicity, cardiotoxicity, and embryonic abnormalities in marine organisms (Honda & Suzuki 2020). Environmental and health concerns associated with PAHs have led to increased efforts to monitor and regulate their presence. Strategies to reduce PAH exposure include controlling emissions from industrial processes, improving combustion efficiency, and employing proper waste disposal methods.

### Loading Packages
```{r load-packages}
library(tidyverse)
library(sf)
library(leaflet)
library(ggplot2)
library(RColorBrewer)
#install.packages("ggridges")
library(ggridges)
```

### Loading PAH Dataset
```{r load-data}
PAHs <- read.csv(paste0("/cloud/project/data/datasets_csv/PAHs_loc.csv"), header = T)
```

### Glimpsing the PAH Dataset
```{r glimpse-data, warning=FALSE}
glimpse(PAHs)
```

## Data Transformation
### Pivoting PAHs into a Longer Format 
This function creates two columns to properly contain the values of PAHs measured during these surveys - the `pah_detected` column contains the type of PAH that was measured and the `amount_detected` column contains the values for each measurement. Selecting specific variables within the `PAHs` data set will allow for a more concise data frame that is more applicable to the analysis.

```{r pivoting-pah, warning=FALSE}
PAHs <- PAHs %>%
  mutate(CHRYSENE = as.numeric(CHRYSENE)) %>%
  mutate(SAMP_DATE1 = as.Date(SAMP_DATE1)) %>%
  pivot_longer(cols = `BENZNE`:`PERYLENE`, 
               names_to = "pah_detected", 
               values_to = "amount_detected")

PAHs = select(PAHs, c(UNIQUE_ID, LATITUDE, LONGITUDE, SOUNDING_M, QUAD_NAME, GEN_LOC_NM, SPECFC_LOC, STATE_NAME, AREA_CODE, SAMP_DATE1, TO_SMP_DT2, DPTH_N_COR, DPTH_CODE, COR_GRB_CD, site, pah_detected, amount_detected))

glimpse(PAHs)
```

## Data Visualizations
### Descriptive Visualizations
First, the number of observations collected at each location will be visualized. 
By counting the observations at each general location, this graph reflects sampling intensity. The greatest number of observations are in the Gulf of Maine below the 50m isobath (as a product of area and sampling intensity); Massachusetts Bays, and Boston Harbor sites. The MA/NH/ME coast are all generally heavily sampled.

``` {r descriptive-vis-pah-locations}
PAHs %>%
  ggplot(aes(x = fct_infreq(GEN_LOC_NM), fill = fct_infreq(GEN_LOC_NM))) +
  geom_bar(stat = "count", color = "black") +
  coord_flip() +
  theme_minimal() +
  scale_fill_viridis_d() +
  theme(legend.position = "none") +
  labs(title = "Number of Observations by General Location",
       subtitle = "for PAH Contaminents",
       x = "General Location",
       y = "Count (n Observations)")
```

## General PAH Concentration
As shown in the plot below, there is a high concentration of Fluoranthene, followed by Pyrene, Chrysene, and Phenanthrene.

``` {r plotting-pahs-detected, warning=FALSE}
PAHs %>%
  ggplot(aes(x = fct_rev(fct_reorder(pah_detected, amount_detected)), y = amount_detected, fill = fct_infreq(pah_detected)))+
  geom_col()+
  scale_fill_viridis_d() +
  theme_bw()+
  coord_flip()+
  theme(legend.position = "none")+
  labs(title = "Contaminents Detected in Gulf of Maine Sediments",
       subtitle = "by PAH concentration",
       x = "PAH Detected",
       y = "Amount Detected")
```

### PAH Concentration in the Gulf of Maine
Of the entire Gulf of Maine, the PAH with the highest average concentration was Fluoranthene (1962.25 ng/g). This is followed by Pyrene (934.50 ng/g), Chrysene (887.80 ng/g), and Phenanthrene (546.57 ng/g). 

AP - *add interpretation here that talks about the properties of each PAH?*


``` {r pah-concentration-histogram}
PAHs %>%
  filter(amount_detected <= 100 ) %>%
  filter(amount_detected >= 1) %>%
  ggplot(aes(x = amount_detected, y = pah_detected, fill = pah_detected)) + 
  geom_density_ridges(alpha=0.6, bins=100) +
  theme_ridges() + 
  theme(legend.position = "none", panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8))+
  theme_minimal() +
  scale_fill_viridis_d() +
  labs(title = "PAH Concentration (ng/g)",
       subtitle = "Fluoranthene, Pyrene, Chrysene, Phenanthrene", 
        x = "Amount Detected (ng/g)", 
        y = "Frequency of Amount Detected", 
       fill = "PAH Detected")

PAHs %>%
  group_by(pah_detected) %>%
  filter(GEN_LOC_NM %in%c("Rockland to north")) %>%
  summarise(amount_detected = mean(amount_detected, na.rm = TRUE), 
            sd_PAH = sd(amount_detected, na.rm = TRUE)) %>%
  arrange(desc(amount_detected))
```
```

# PAH Site to Site Comparisons
## General PAH Sumary Statistics
The table below is visualizing the mean, standard deviation, count, and principal square root values of PAH concentrations for the general locations. 
Massachusetts Bays has the highest count of PAHs with 3,105 observation, followed by Northwest Boston Harbor (1,820 observations) and Cape Elizabeth to Rockland (860 observations). This does not suggest that these locations will have the highest concentrations of PAHs, as this does not reflect the quantity of PAHs collected. Boston Inner Harbor (2,291.16 ng/g) has the highest mean PAH concentration, followed by Northwest Boston Harbor (659.17 ng/g) and Central Boston Harbor (633.28 ng/g). In addition to this, Boston Inner Harbor has the highest standard deviation (15,446.77), followed by Northwest Boston Harbor (3,177.42) and Cape Elizabeth to Rockland (2,514.69). 

``` {r summary-stats-PAHs}
Summary_PAHs <- PAHs %>%
  group_by(GEN_LOC_NM) %>%
  drop_na(amount_detected) %>%
  summarise(mean_PAH = mean(amount_detected),
    sd_PAH = sd(amount_detected),
    n_PAH = n(),
    SE_PAH = sd(amount_detected) / sqrt(n())) 

Summary_PAHs
```

AP- I'm not sure where to put the graph below.
This plot shows us the exact quantity of PAH found at each site. 

```{r summarize-by-site}
PAHs %>%
  group_by(GEN_LOC_NM, pah_detected) %>%
  summarise(amount_detected = sum(amount_detected, na.rm = TRUE)) %>%
  arrange(desc(amount_detected))
```

The graph below is showing which site has the highest mean PAH concentrations in the Gulf of Maine. As was reflected above, the site with the highest PAH concentration is Boston Inner Harbor, followed by Northwest Boston Harbor, Central Boston Harbor, and Cape Elizabeth to Rockland, ME. 

```{r pah-gen-loc}
Summary_PAHs %>%
  ggplot(aes(x = fct_rev(fct_reorder(GEN_LOC_NM, mean_PAH)), y = mean_PAH, fill = GEN_LOC_NM))+
  geom_col(col = "black")+
  geom_errorbar(aes(ymin = mean_PAH - SE_PAH, ymax = mean_PAH + SE_PAH), width = 0.2)+
  coord_flip()+
  scale_fill_viridis_d()+
  theme_bw()+
  theme(legend.position = "none")+
  labs(title = "PAH Concentrations in Gulf of Maine Sediments",
       x = "General location",
       y = "Mean total PAH concentration ng/g",
       caption = "Error bars = 1 standard error")
```

# Visualizing Specific PAH Concentrations
## PCB Summary Statistics of Harbors With Highest Concentrations
This graph looks specifically at the locations with the highest concentrations of PCBs.

``` {r}
Summary_PAHs %>%
  filter(GEN_LOC_NM %in% c("Boston Inner Harbor", "Northwest Boston Harbor", "Central Boston Harbor", "Cape Elizabeth to Rockland", "Cape Ann to Cape Elizabeth")) %>%
  ggplot(aes(x = fct_rev(fct_reorder(GEN_LOC_NM, mean_PAH)), y = mean_PAH, fill = GEN_LOC_NM)) +
  geom_col() +
  geom_errorbar(aes(ymin = mean_PAH - SE_PAH, ymax = mean_PAH + SE_PAH), width = 0.2) +
  scale_fill_brewer(type = "qual", palette = 4, direction = 1, aesthetics = "fill") +
  theme_minimal() +
  scale_fill_viridis_d()+
  coord_flip() +
  theme(legend.position = "none") +
  labs(title = "Locations with the highest PAH Concentrations",
       x = "General location",
       y = "Mean total PCB concentration ng/g",
       caption = "Error bars = 1 standard error")
```

### Boston Harbor (General) Concentrations
This plot visualizes the specific PAHs present in Boston Harbor. 
As shown in the plot, Chrysene (2234.89 ng/g) has the highest average concentration in the harbor. This is followed by Pyrene (average of 2197.45 ng/g), Fluoranthene (average of 2016.27 ng/g), and Phenanthrene (average of 1307.48 ng/g). 

```{r sum-pah-boston-general-harbor, message=FALSE, echo=FALSE}
PAHs %>%
  filter(GEN_LOC_NM %in%c("Boston Inner Harbor", "Northwest Boston Harbor", "Central Boston Harbor", "Southeast Boston Harbor")) %>%
  group_by(pah_detected) %>%
  summarise(amount_detected = mean(amount_detected, na.rm = TRUE)) %>%
  arrange(desc(amount_detected))
```

``` {r pahs-boston-harbor-graph}
PAHs %>%
  group_by(GEN_LOC_NM, pah_detected) %>%
  summarise(amount_detected = sum(amount_detected, na.rm = TRUE)) %>%
  filter(GEN_LOC_NM %in% c("Boston Inner Harbor", "Northwest Boston Harbor", "Central Boston Harbor", "Southeast Boston Harbor")) %>%
  filter(pah_detected %in% c("PYRENE", "FLORNTHN", "PHNANTHR", "CHRYSENE", "ANTHRACN", "PERYLENE", "FLUORENE", "ACNPHTHN")) %>%
ggplot(aes(x = fct_rev(fct_reorder(pah_detected, amount_detected)), y = amount_detected, fill = pah_detected))+
  geom_col()+
  coord_flip()+
  theme_minimal() +
  scale_fill_brewer(type = "qual", palette = 8, direction = 1, aesthetics = "fill") +
  theme(legend.position = "none")+
  labs(title = "PAH Concentrations in the Gulf of Maine",
       subtitle = "in Boston Harbor",
       x = "",
       y = "Percentage Detected", 
       fill = "PAH Detected")
```

``` {r pie-chart-boston}
PAHs %>%
  group_by(GEN_LOC_NM, pah_detected) %>%
  summarise(amount_detected = sum(amount_detected, na.rm = TRUE)) %>%
  filter(GEN_LOC_NM %in% c("Boston Inner Harbor", "Northwest Boston Harbor", "Central Boston Harbor", "Southeast Boston Harbor")) %>%
  filter(pah_detected %in% c("PYRENE", "FLORNTHN", "PHNANTHR", "CHRYSENE", "ANTHRACN", "PERYLENE", "FLUORENE", "ACNPHTHN")) %>%
ggplot(aes(x="", y= amount_detected, fill = pah_detected)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
    theme_minimal() +
  scale_fill_brewer(type = "qual", palette = 8, direction = 1, aesthetics = "fill") +
  labs(title = "PAH Concentrations in the Gulf of Maine",
       subtitle = "in Boston Harbor",
       x = "",
       y = "Percentage Detected", 
       fill = "PAH Detected")
``` 

### Boston Inner Harbor Concentrations
This plot visualizes the specific PAHs present in Boston Inner Harbor. 
As shown in the plot, Chrysene (569,679.08 ng/g) has the highest concentration in the harbor. This is followed by Pyrene (346679.56 ng/g), Fluoranthene (319594.32 ng/g), and Phenanthrene (221,665.68 ng/g). 

```{r sum-organics-boston-inner-harbor, message=FALSE, echo=FALSE}
PAHs %>%
  group_by(GEN_LOC_NM, pah_detected) %>%
  summarise(amount_detected = sum(amount_detected, na.rm = TRUE)) %>%
  filter(GEN_LOC_NM == "Boston Inner Harbor") %>%
  arrange(desc(amount_detected))

``` 
``` {r sum-organics-boston-inner-harbor-graph}
PAHs %>%
  group_by(GEN_LOC_NM, pah_detected) %>%
  summarise(amount_detected = sum(amount_detected, na.rm = TRUE)) %>%
  filter(GEN_LOC_NM == "Boston Inner Harbor") %>%
 ggplot(aes(x = fct_rev(fct_reorder(pah_detected, amount_detected)), y = amount_detected, fill = pah_detected))+
  geom_col()+
  coord_flip()+
  theme_minimal() +
  theme(legend.position = "none")+
  labs(title = "PAH Concentrations in the Gulf of Maine",
       subtitle = "in Boston Inner Harbor",
       x = "PAH Detected",
       y = "Amount Detected (ng/g)")
```

### Northwest Boston Harbor Concentrations
This plot visualizes the specific PAHs present in Northwest Boston Harbor.  
As shown in the plot, Pyrene (315,904.49 ng/g) has the highest concentration in the harbor. This is followed by Fluoranthene (312,942.98 ng/g), Phenanthrene (219,061.23 ng/g), and Chrysene (200,784.37 ng/g).

```{r type_of_organic-in-northwest-boston-harbor, echo=FALSE, message=FALSE}
PAHs %>%
  group_by(GEN_LOC_NM, pah_detected) %>%
  summarise(amount_detected = sum(amount_detected, na.rm = TRUE)) %>%
  filter(GEN_LOC_NM == "Northwest Boston Harbor") %>%
  arrange(desc(amount_detected))
```
``` {r type_of_organic-in-northwest-boston-harbor-graph, warning = FALSE}
PAHs %>%
  drop_na(amount_detected) %>%
  filter(GEN_LOC_NM %in% c("Northwest Boston Harbor")) %>%
  ggplot(aes(x = fct_rev(fct_reorder(pah_detected, amount_detected)), y = amount_detected, fill = pah_detected))+
  geom_col()+
  coord_flip()+
  theme(legend.position = "none")+
  labs(title = "PAH Concentrations in Northwest Boston Harbor",
       x = "PAH Detected",
       y = "Amount Detected (ng/g)")
```

## Central Boston Harbor Concentrations
This plot visualizes the specific PAHs present from Central Boston Harbor.  
As shown in the plot, Pyrene (71,710.28 ng/g) has the highest concentration in the harbor. This is followed by Fluoranthene (64,217.22 ng/g), and Phenanthrene (43304.62 ng/g). 
``` {r sum-organics-boston-central-harbor, echo=FALSE, message=FALSE}
PAHs %>%
  group_by(GEN_LOC_NM, pah_detected) %>%
  summarise(amount_detected = sum(amount_detected, na.rm = TRUE)) %>%
  filter(GEN_LOC_NM == "Central Boston Harbor") %>%
  arrange(desc(amount_detected))
```
```{r sum-organics-boston-central-harbor-graph, warning=FALSE}
PAHs %>%
  drop_na(amount_detected) %>%
  filter(GEN_LOC_NM %in% c("Central Boston Harbor")) %>%
 ggplot(aes(x = fct_rev(fct_reorder(pah_detected, amount_detected)), y = amount_detected, fill = pah_detected))+
  geom_col()+
  coord_flip()+
  theme(legend.position = "none")+
  labs(title = "PAH Concentrations in Central Boston Harbor",
       x = "PAH Detected",
       y = "Amount Detected (ng/g)")
```

## Cape Ann to Cape Elizabeth Concentrations
This plot visualizes the specific PAHs present from Cape Ann to Cape Elizabeth.  
As shown in the plot, Fluoranthene (77,701.85 ng/g) has the highest concentration in the harbor. This is followed by Pyrene	(73,067.06 ng/g), Phenanthrene (46,915.83 ng/g), and Chrysene (35,656.56 ng/g). 

```{r type_of_organic-cape-ann-to-elizabeth, echo=FALSE, message=FALSE}
PAHs %>%
  group_by(GEN_LOC_NM, pah_detected) %>%
  summarise(amount_detected = sum(amount_detected, na.rm = TRUE)) %>%
  filter(GEN_LOC_NM == "Cape Ann to Cape Elizabeth") %>%
  arrange(desc(amount_detected))
```
```{r type_of_organic-cape-ann-to-elizabeth-graph, warning=FALSE}
PAHs %>%
  drop_na(amount_detected) %>%
  filter(GEN_LOC_NM %in% c("Cape Ann to Cape Elizabeth")) %>%
 ggplot(aes(x = fct_rev(fct_reorder(pah_detected, amount_detected)), y = amount_detected, fill = pah_detected))+
  geom_col()+
  coord_flip()+
  theme_minimal() +
  scale_fill_viridis_d() +
  theme(legend.position = "none")+
  labs(title = "PAH Concentrations from Cape Ann to Cape Elizabeth",
       x = "PAH Detected",
       y = "Amount Detected (ng/g)")
```

## Cape Elizabeth to Rockland Concentrations
This plot visualizes the specific PAHs present from Cape Elizabeth to Rockland.  
As shown in the plot, Fluoranthene (116,706.129 ng/g) has the highest concentration in the harbor. This is followed by Chrysene (112,928.05 ng/g), Pyrene (110,687.928 ng/g), and Phenanthrene (40,140.04 ng/g). 

```{r type_of_organic-in-cape-elizabeth-to-rockland, echo=FALSE, message=FALSE}
PAHs %>%
  group_by(GEN_LOC_NM, pah_detected) %>%
  summarise(amount_detected = sum(amount_detected, na.rm = TRUE)) %>%
  filter(GEN_LOC_NM == "Cape Elizabeth to Rockland") %>%
  arrange(desc(amount_detected))
```
```{r type_of_organic-in-cape-elizabeth-to-rockland-graph, warning=FALSE}
PAHs %>%
  drop_na(amount_detected) %>%
  filter(GEN_LOC_NM %in% c("Cape Elizabeth to Rockland")) %>%
 ggplot(aes(x = fct_rev(fct_reorder(pah_detected, amount_detected)), y = amount_detected, fill = pah_detected))+
  geom_col()+
  coord_flip()+
  theme_minimal() +
  scale_fill_viridis_d() +
  theme(legend.position = "none")+
  labs(title = "PAH Concentrations from Cape Elizabeth to Rockland",
       x = "PAH Detected",
       y = "Amount Detected (ng/g)")

PAHs %>%
  drop_na(amount_detected) %>%
  filter(GEN_LOC_NM %in% c("Cape Elizabeth to Rockland")) %>%
ggplot(aes(x="", y = amount_detected, fill = pah_detected)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0)
```

### Boston Harbor (General) Concentrations
This plot visualizes the specific PAHs present in Boston Harbor. 
As shown in the plot, Chrysene (824,674.86 ng/g) has the highest concentration in the harbor. This is followed by Pyrene (753,726.83 ng/g), Fluoranthene (721,826.05 ng/g), and Phenanthrene (494,229.22 ng/g). 

```{r sum-pah-boston-general-harbor-2, message=FALSE, echo=FALSE}
PAHs %>%
  filter(GEN_LOC_NM %in%c("Rockland to north")) %>%
  group_by(pah_detected) %>%
  summarise(amount_detected = sum(amount_detected, na.rm = TRUE)) %>%
  arrange(desc(amount_detected))
```

``` {r pahs-boston-harbor-graph-2}
PAHs %>%
  group_by(GEN_LOC_NM, pah_detected) %>%
  summarise(amount_detected = sum(amount_detected, na.rm = TRUE)) %>%
  filter(GEN_LOC_NM %in% c("Rockland to north")) %>%
  filter(pah_detected %in% c("PYRENE", "FLORNTHN", "PHNANTHR", "CHRYSENE", "ANTHRACN", "PERYLENE", "FLUORENE", "ACNPHTHN")) %>%
ggplot(aes(x = fct_rev(fct_reorder(pah_detected, amount_detected)), y = amount_detected, fill = pah_detected))+
  geom_col()+
  coord_flip()+
  theme_minimal() +
  scale_fill_brewer(type = "qual", palette = 8, direction = 1, aesthetics = "fill") +
  theme(legend.position = "none")+
  labs(title = "PAH Concentrations in the Gulf of Maine",
       subtitle = "of MDI and Penobscot, Maine",
       x = "",
       y = "Percentage Detected", 
       fill = "PAH Detected")
```

``` {r pie-chart-boston}
PAHs %>%
  group_by(GEN_LOC_NM, pah_detected) %>%
  summarise(amount_detected = sum(amount_detected, na.rm = TRUE)) %>%
  filter(GEN_LOC_NM %in% c("Rockland to north")) %>%
  filter(pah_detected %in% c("PYRENE", "FLORNTHN", "PHNANTHR", "CHRYSENE", "ANTHRACN", "PERYLENE", "FLUORENE", "ACNPHTHN")) %>%
ggplot(aes(x="", y= amount_detected, fill = pah_detected)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
    theme_minimal() +
  scale_fill_brewer(type = "qual", palette = 8, direction = 1, aesthetics = "fill") +
  labs(title = "PAH Concentrations in the Gulf of Maine",
       subtitle = "of MDI and Penobscot, Maine",
       x = "",
       y = "Percentage Detected", 
       fill = "PAH Detected")
``` 

### PAH Summary Statistics of Rivers With Highest Concentrations
As stated in our project proposal, we were interested in visualizing contaminants in Maine Rivers as well as harbors and bays. The table below is visualizing the mean, standard deviation, count, and principal square root values of PAH concentrations for the general river locations.

```{r calculate-sum-pah-by-site}
Sum_PAH_site <- PAHs %>%
  group_by(site) %>%
  drop_na(amount_detected) %>%
  summarise(mean_PAH_T = mean(amount_detected),
    sd_PAH_T = sd(amount_detected),
    n_PAH_T = n(),
    SE_PAH_T = sd(amount_detected) / sqrt(n()))

Sum_PAH_site
```

Below we are identifying specific Maine rivers to use in our analysis.

```{r identify-ME-rivers, message=FALSE}
PAHs %>%
  filter(STATE_NAME == "ME") %>%
  filter(amount_detected != "NA") %>%
  distinct(site)
```

Fore River, Royal River, Kennebunk River, York River

Below we are visualizing the mean PAHs present in Maine rivers. 

```{r bar-pah-me-rivers-without-union}
Sum_PAH_site %>%
  filter(site %in% c("York River", "Kennebunk River", "Fore River", "Royal River")) %>%
  ggplot(aes(x = fct_rev(fct_reorder(site, mean_PAH_T)), y = sapply(mean_PAH_T, FUN=function(x) ifelse(x==0.000000e0, -0.2,x)), fill = fct_rev(fct_reorder(site, mean_PAH_T)))) +
  geom_bar(stat="identity", col = "black") +
  scale_x_discrete(drop=FALSE) +
  geom_errorbar(aes(ymin = mean_PAH_T - SE_PAH_T, ymax = mean_PAH_T + SE_PAH_T), width = 0.2) +
  scale_fill_manual(values = c("Royal River" = "green4",
                               "York River" = "darkgoldenrod4",
                               "Kennebunk River" = "firebrick",
                               "Fore River" = "peachpuff")) +
  theme_bw() +
  coord_flip() +
  theme(legend.position = "none") +
  labs(title = "PAH Concentrations in Maine Rivers",
       subtitle = "Union River excluded",
       caption = "Error bars = 1 standard error",
       x = "Location",
       y = "Mean total PAH concentration ug/g")
```

Next, we will analyze Rivers in Massachusetts. 

```{r identiy-ma-rivers, message=FALSE}
PAHs %>%
  filter(STATE_NAME == "MA") %>%
  filter(pah_detected != "NA") %>%
  distinct(site, pah_detected, amount_detected) %>%
  arrange(desc(amount_detected))
```

```{r bar-pah-mass-rivers}
Sum_PAH_site %>%
  filter(site %in% c("Chelsea River", "Mystic River", "Fore River", "Charles River")) %>%
  ggplot(aes(x = fct_rev(fct_reorder(site, mean_PAH_T)), y = sapply(mean_PAH_T, FUN=function(x) ifelse(x==0.000000e0, -0.2,x)), fill = fct_rev(fct_reorder(site, mean_PAH_T)))) +
  geom_bar(stat="identity", col = "black") +
  scale_x_discrete(drop=FALSE) +
  geom_errorbar(aes(ymin = mean_PAH_T - SE_PAH_T, ymax = mean_PAH_T + SE_PAH_T), width = 0.2) +
  scale_fill_manual(values = c("Chelsea River" = "green4",
                               "Mystic River" = "darkgoldenrod4",
                               "Charles River" = "firebrick",
                               "Fore River" = "peachpuff")) +
  theme_bw() +
  coord_flip() +
  theme(legend.position = "none") +
  labs(title = "PAH Concentrations in Massachusetts Rivers",
       caption = "Error bars = 1 standard error",
       x = "Location",
       y = "Mean total PAH concentration ug/g")
```

## Statistical tests
### Difference between general locations

Does total PAH concentration differ significantly between general locations?

```{r}
hist(PAHs$amount_detected)
```

```{r}
kw1=kruskal.test(PAHs$amount_detected ~ PAHs$GEN_LOC_NM)
kw1
```

This shows that it is significant. We will now run a Dunn post-hoc test to identify any significance between locations.

```{r}
#install.packages("dunn.test")
#library(dunn.test)

#dunn=dunn.test(PAHs$amount_detected,Organics$GEN_LOC_NM,method="bh")

```

Interpretation: Most general locations have significantly different mean total PAH concentrations.

## Map Plots
### Static maps

```{r load-GOM_states-shp}
GOM_states <- st_read("/cloud/project/extra/GOM_DD.shp")
```

```{r load-bathy-shp}
Bathy <- st_read("/cloud/project/extra/BATHYMGM_ARC.shp")
```

Bathymetry data is in projected coordinate system (NAD83), other data is geodetic (NAD83).  
Wasn't able to add this layer to map plot -- need to change projection.

```{r bathy-to-NAD83-format}
st_crs(Bathy)
Bathy <- st_transform(Bathy, "+init=epsg:4269")
```

Now the Bathy shapefile is in geodetic NAD83 format.

```{r unique-bathy}
unique(Bathy$CONTOUR)
```

```{r plot-bathy}
Bathy_low_res <- Bathy%>%
 filter(CONTOUR %in% c("-100","-500","-1000","-2000","-3000","-4000"))
ggplot(Bathy_low_res) +
  geom_sf(aes())
```
```{r creating-Bathy_hi_res-pah}
Bathy_hi_res <- Bathy%>%
 filter(CONTOUR %in% c("-40","-80","-120","160","200","240"))
ggplot(Bathy_hi_res) +
  geom_sf(aes())
```

```{r drop-na-pah}
PAH_no_na <- PAHs %>%
  drop_na(amount_detected)

PAH_no_na_specific_pah <- PAH_no_na %>%
  filter(pah_detected %in% c("BENZNE", "BENZNE", "NAPHTHLN", "BIPHENYL", "ACNPHTHN", "ACNPHTYL", "FLUORENE", "PHNANTHR", "ANTHRACN", "PYRENE", "FLORNTHN", "CHRYSENE", "PERYLENE"))

```

```{r pah-gom-map-plot, warning = FALSE}
ggplot(GOM_states)+
  geom_sf(aes())+
  geom_sf(data = Bathy_low_res, color = "gray80", width = 1)+
  geom_point(data = PAH_no_na, (aes(x = LONGITUDE, y = LATITUDE, size = amount_detected, alpha = 0.5)))+
  xlim(-72,-65)+
  ylim(40,45)+
  theme_bw()+
  labs(title = "Distribution and concentration of PAHs",
       subtitle ="Gulf of Maine sediments",
       x = "Longitude",
       y = "Latitude")+
  guides(size = guide_legend(title = "PAH ng/g"))+
  guides(alpha = FALSE)+
  ggspatial::annotation_scale(
    location = "bl",
    bar_cols = c("grey60", "white"),
    text_family = "ArcherPro Book"
  ) +
  ggspatial::annotation_north_arrow(
    location = "tr", which_north = "true",
    pad_x = unit(0, "in"), pad_y = unit(0.2, "in"),
    style = ggspatial::north_arrow_nautical(
      fill = c("grey40", "white"),
      line_col = "grey20",
      text_family = "ArcherPro Book"))
```

Above is the Distribution and concentration of PAHs in the Gulf of Maine sediments. 

### MDI Area Mapped
```{r MDI-area-map-pah}
ggplot(GOM_states)+
  geom_sf(aes())+
  geom_sf(data = Bathy_hi_res, color = "gray80", width = 1)+
  geom_point(data=PAH_no_na, (aes(x = LONGITUDE, y = LATITUDE, size = amount_detected, alpha = 0.5)))+
  xlim(-69.2,-68)+
  ylim(44,44.5)+
  theme_bw()+
  labs(title = "Distribution and concentration of PAHs",
       subtitle ="MDI and Penobscot Bay sediments",
       x = "Longitude",
       y = "Latitude")+
  guides(size = guide_legend(title = "PAH ng/g"))+
  guides(alpha = FALSE)+
  ggspatial::annotation_scale(
    location = "bl",
    bar_cols = c("grey60", "white"),
    text_family = "ArcherPro Book"
  ) +
  ggspatial::annotation_north_arrow(
    location = "tr", which_north = "true",
    pad_x = unit(0, "in"), pad_y = unit(0.2, "in"),
    style = ggspatial::north_arrow_nautical(
      fill = c("grey40", "white"),
      line_col = "grey20",
      text_family = "ArcherPro Book"))
```

```{r MDI-area-map-pah}
ggplot(GOM_states)+
  geom_sf(aes())+
  geom_sf(data = Bathy_hi_res, color = "gray80", width = 1)+
  geom_point(data=PAH_no_na_specific_pah, (aes(x = LONGITUDE, y = LATITUDE, scale = amount_detected, color = pah_detected, alpha = 0.5)))+
  xlim(-69.2,-68)+
  ylim(44,44.5)+
  theme_bw()+
  labs(title = "Distribution and concentration of PAHs",
       subtitle ="MDI and Penobscot Bay sediments",
       x = "Longitude",
       y = "Latitude", 
       color = "PAH Detected")+
  guides(size = guide_legend(title = "PAH ng/g"))+
  guides(alpha = FALSE)+
  ggspatial::annotation_scale(
    location = "bl",
    bar_cols = c("grey60", "white"),
    text_family = "ArcherPro Book"
  ) +
  ggspatial::annotation_north_arrow(
    location = "tr", which_north = "true",
    pad_x = unit(0, "in"), pad_y = unit(0.2, "in"),
    style = ggspatial::north_arrow_nautical(
      fill = c("grey40", "white"),
      line_col = "grey20",
      text_family = "ArcherPro Book"))

```

Above is the Distribution and concentration of PAHs in MDI and Penobscot Bay sediments. 

### Boston, MA Area Mapped
```{r boston-and-ma-area-map-pah}
ggplot(GOM_states)+
  geom_sf(aes())+
  geom_sf(data = Bathy_hi_res, color = "gray80", width = 1)+
  geom_point(data=PAH_no_na, (aes(x = LONGITUDE, y = LATITUDE, size = amount_detected, color = pah_detected, alpha = 0.5)))+
  xlim(-71.2,-69.5)+
  ylim(41.8,43)+
  theme_bw()+
  labs(title = "Distribution and concentration of PAHs",
       subtitle ="Boston, MA and Cape Cod Bay sediments",
       x = "Longitude",
       y = "Latitude")+
  guides(size = guide_legend(title = "PAH ng/g"))+
  guides(alpha = FALSE)+
  ggspatial::annotation_scale(
    location = "bl",
    bar_cols = c("grey60", "white"),
    text_family = "ArcherPro Book"
  ) +
  ggspatial::annotation_north_arrow(
    location = "tr", which_north = "true",
    pad_x = unit(0, "in"), pad_y = unit(0.1, "in"),
    style = ggspatial::north_arrow_nautical(
      fill = c("grey40", "white"),
      line_col = "grey20",
      text_family = "ArcherPro Book"))
```

Above is the Distribution and concentration of PAHs in Boston, MA and Cape Cod Bay sediments. 

```{r boston-and-ma-area-map-pah-colored}
ggplot(GOM_states)+
  geom_sf(aes())+
  geom_sf(data = Bathy_hi_res, color = "gray80", width = 1)+
  geom_point(data=PAH_no_na_specific_pah, (aes(x = LONGITUDE, y = LATITUDE, size = amount_detected, color = pah_detected, alpha = 0.5)))+
  xlim(-71.2,-69.5)+
  ylim(41.8,43)+
  theme_bw()+
  labs(title = "Distribution and concentration of PAHs",
       subtitle ="Boston, MA and Cape Cod Bay sediments",
       x = "Longitude",
       y = "Latitude", 
       color = "PAH Detected")+
  guides(size = guide_legend(title = "PAH ng/g"))+
  guides(alpha = FALSE)+
  ggspatial::annotation_scale(
    location = "bl",
    bar_cols = c("grey60", "white"),
    text_family = "ArcherPro Book"
  ) +
  ggspatial::annotation_north_arrow(
    location = "tr", which_north = "true",
    pad_x = unit(0, "in"), pad_y = unit(0.1, "in"),
    style = ggspatial::north_arrow_nautical(
      fill = c("grey40", "white"),
      line_col = "grey20",
      text_family = "ArcherPro Book"))
```

Above is the same graph, but now we are coloring the points by the specific pah that was detected. 

### Interactive map

```{r create-html-labels-pah}
#labels <- sprintf("<strong>%s</strong><br/>%g ng/g", 
#                 PAH_no_na$SPECFC_LOC, PAH_no_na$amount_detected) %>%
# lapply(htmltools::HTML)
#
#head(labels, 1)
```

```{r leaflet-map-pcb} 
#leaflet(data = PAH_no_na) %>%
  #addProviderTiles(providers$Esri.WorldTopoMap) %>%
  #setView(lng = -68.5, 
    #      lat = 43.5, 
     #     zoom = 6) %>%
  #addCircleMarkers(lng = ~LONGITUDE, lat = ~LATITUDE, popup = c(~SPECFC_LOC, ~amount_detected), label = labels)
```

Note that the two above code chunks are commented out as .rmd will not knit to github document with html functions.  Uncomment to run and change output type to `html_document` to knit.


