---
title: 'GOM Contaminated Sediments Analysis: PAHs'
author: "Joshua Harkness and Autumn Pauly"
date: "2023-10-28"
output: github_document
---
# Introduction
This document is an analysis of distributions and concentrations of Polycyclic aromatic hydrocarbons (PAH) contaminants that were reported in a publication by the U.S. Geological Survey on contaminated sediments in the Gulf of Maine. In this data set, we analyzed the prevalence of Benzene, Dibenzothiophene, Dibenzofuran, Biphenyl, Acenaphthene, Acenaphthylene, Fluorene, Phenanthrene, Anthracene,  Pyrene , Fluoranthene, Chrysene, and Perylene. 

Polycyclic aromatic hydrocarbons (PAH) are a class of organic compound made of multiple aromatic rings. Particle-bound PAHs, or PAHs with five rings or more, are considered to be very hazardous to human health. These are the types of chemicals that naturally occur in crude oil, coal, and gasoline and are put into the environment in high temperature events, such as burning, and oil spill events. 

The Agency for Toxic Substances and Disease Registry states that PAHs have a low degree of acute toxicity to humans. When chronically exposed as high doses, PAH toxicity can induce issues within the pulmonary, gastrointestinal, renal, and dental systems as well as increase an individual's risk of cancer (Gupta et al. 1991). In regards to the environment, PAHs have a moderate to high acute toxicity to wildlife, specifically aquatic life and birds (Honda & Suzuki 2020). Research has found that PAHs have toxic compounds that affect immunotoxicity, cardiotoxicity, and embryonic abnormalities in marine organisms.


### Loading Packages
```{r load-packages}
library(tidyverse)
library(sf)
library(leaflet)
library(ggplot2)
library(RColorBrewer)
install.packages("hrbrthemes")
library(hrbrthemes)
```

### Loading PAH Dataset
```{r load-data}
PAHs <- read.csv(paste0("/cloud/project/data/datasets_csv/PAHs_loc.csv"), header = T)
```

### Glimpsing the PAH Dataset
```{r glimpse-data, warning=FALSE}
glimpse(PAHs)
```

## Data Transformation
### Pivoting PAHs into a Longer Format 
This function creates two columns to properly contain the values of PAHs measured during these surveys - the `pah_detected` column contains the type of PAH that was measured and the `amount_detected` column contains the values for each measurement. Selecting specific variables within the `PAHs` data set will allow for a more concise data frame that is more applicable to the analysis.
```{r pivoting-pah, warning=FALSE}
PAHs <- PAHs %>%
  mutate(CHRYSENE = as.numeric(CHRYSENE)) %>%
  pivot_longer(cols = `BENZNE`:`PERYLENE`, 
               names_to = "pah_detected", 
               values_to = "amount_detected")

PAHs = select(PAHs, c(UNIQUE_ID, LATITUDE, LONGITUDE, SOUNDING_M, QUAD_NAME, GEN_LOC_NM, SPECFC_LOC, AREA_CODE, SAMP_DATE1, TO_SMP_DT2, DPTH_N_COR, DPTH_CODE, COR_GRB_CD, site, pah_detected, amount_detected))

glimpse(PAHs)
```

## Data Visualizations
### Descriptive Visualizations
First, the number of observations collected at each location will be visualized. 
By counting the observations at each general location, this graph reflects sampling intensity. The greatest number of observations are in the Gulf of Maine below the 50m isobath (as a product of area and sampling intensity); Massachusetts Bays, and Boston Harbor sites. The MA/NH/ME coast are all generally heavily sampled.

``` {r descriptive-vis-pah-locations}
PAHs %>%
  ggplot(aes(x = fct_infreq(GEN_LOC_NM), fill = fct_infreq(GEN_LOC_NM))) +
  geom_bar(stat = "count", color = "black") +
  coord_flip() +
  theme_minimal() +
  scale_fill_viridis_d() +
  theme(legend.position = "none") +
  labs(title = "Number of Observations by General Location",
       subtitle = "for PAH Contaminents",
       x = "General Location",
       y = "Count (n Observations)")
```

## General PAH Concentration
As shown in the plot below, there is a high concentration of Fluoranthene, followed by Pyrene, Chrysene, and Phenanthrene.

``` {r plotting-pahs-detected, warning=FALSE}
PAHs %>%
  ggplot(aes(x = fct_rev(fct_reorder(pah_detected, amount_detected)), y = amount_detected, fill = fct_infreq(pah_detected)))+
  geom_col()+
  scale_fill_viridis_d() +
  theme_bw()+
  coord_flip()+
  theme(legend.position = "none")+
  labs(title = "Contaminents Detected in Gulf of Maine Sediments",
       subtitle = "by PAH concentration",
       x = "PAH Detected",
       y = "Amount Detected")
```
### Fluoranthene Concentration

``` {r fluoranthene-concentration}
PAHs %>%
  filter(pah_detected == "FLORNTHN") %>%
  filter(amount_detected <= 4000 ) %>%
  ggplot(aes(x = amount_detected)) + 
  theme_ipsum_ps() +
  geom_histogram(binwidth = 100, fill="#69b3a2", color="#e9ecef", alpha=0.9) + 
  labs(title = "Fluoranthene Concentration (ng/g)", 
        x = "Amount Detected (ng/g)", 
        y = "Frequency of Amount Detected")

```

### Pyrene Concentration

``` {r pyrene-concentration}
PAHs %>%
  filter(pah_detected == "PYRENE") %>%
  filter(amount_detected <= 4000, 
         amount_detected > 0) %>%
  ggplot(aes(x = amount_detected)) + 
  geom_histogram(binwidth = 100, fill="#69b3a2", color="#e9ecef", alpha=0.9) + 
  labs(title = "Pyrene Concentration (ng/g)", 
        x = "Amount Detected (ng/g)", 
        y = "Frequency of Amount Detected")
```

# PAH Site to Site Comparisons
## General PAH Sumary Statistics
The table below is visualizing the mean, standard deviation, count, and principal square root values of PAH concentrations for the general locations. 
Massachusetts Bays has the highest count of PAHs with 3,105 observation, followed by Northwest Boston Harbor (1,820 observations) and Cape Elizabeth to Rockland (860 observations). This does not suggest that these locations will have the highest concentrations of PAHs, as this does not reflect the quantity of PAHs collected. Boston Inner Harbor (2,291.16 ng/g) has the highest mean PAH concentration, followed by Northwest Boston Harbor (659.17 ng/g) and Central Boston Harbor (633.28 ng/g). In addition to this, Boston Inner Harbor has the highest standard deviation (15,446.77), followed by Northwest Boston Harbor (3,177.42) and Cape Elizabeth to Rockland (2,514.69). 

``` {r summary-stats-PAHs}
Summary_PAHs <- PAHs %>%
  group_by(GEN_LOC_NM) %>%
  drop_na(amount_detected) %>%
  summarise(mean_PAH = mean(amount_detected),
    sd_PAH = sd(amount_detected),
    n_PAH = n(),
    SE_PAH = sd(amount_detected) / sqrt(n())) 

Summary_PAHs
```

AP- I'm not sure where to put the graph below.
This plot shows us the exact quantity of PAH found at each site. 
```{r summarize-by-site}
PAHs %>%
  group_by(GEN_LOC_NM, pah_detected) %>%
  summarise(amount_detected = sum(amount_detected, na.rm = TRUE)) %>%
  arrange(desc(amount_detected))
```

The graph below is showing which site has the highest mean PAH concentrations in the Gulf of Maine. As was reflected above, the site with the highest PAH concentration is Boston Inner Harbor, followed by Northwest Boston Harbor, Central Boston Harbor, and Cape Elizabeth to Rockland, ME. 
```{r pah-gen-loc}
Summary_PAHs %>%
  ggplot(aes(x = fct_rev(fct_reorder(GEN_LOC_NM, mean_PAH)), y = mean_PAH, fill = GEN_LOC_NM))+
  geom_col(col = "black")+
  geom_errorbar(aes(ymin = mean_PAH - SE_PAH, ymax = mean_PAH + SE_PAH), width = 0.2)+
  coord_flip()+
  scale_fill_viridis_d()+
  theme_bw()+
  theme(legend.position = "none")+
  labs(title = "PAH Concentrations in Gulf of Maine Sediments",
       x = "General location",
       y = "Mean total PAH concentration ng/g",
       caption = "Error bars = 1 standard error")
```

# Visualizing Specific PAH Concentrations
## PCB Summary Statistics of Harbors With Highest Concentrations
This graph looks specifically at the locations with the highest concentrations of PCBs.

``` {r}
Summary_PAHs %>%
  filter(GEN_LOC_NM %in% c("Boston Inner Harbor", "Northwest Boston Harbor", "Central Boston Harbor", "Cape Elizabeth to Rockland", "Cape Ann to Cape Elizabeth")) %>%
  ggplot(aes(x = fct_rev(fct_reorder(GEN_LOC_NM, mean_PAH)), y = mean_PAH, fill = GEN_LOC_NM)) +
  geom_col() +
  geom_errorbar(aes(ymin = mean_PAH - SE_PAH, ymax = mean_PAH + SE_PAH), width = 0.2) +
  scale_fill_brewer(type = "qual", palette = 4, direction = 1, aesthetics = "fill") +
  theme_minimal() +
  scale_fill_viridis_d()+
  coord_flip() +
  theme(legend.position = "none") +
  labs(title = "Locations with the highest PAH Concentrations",
       x = "General location",
       y = "Mean total PCB concentration ng/g",
       caption = "Error bars = 1 standard error")
```

### Boston Inner Harbor Concentrations
This plot visualizes the specific PAHs present in Boston Inner Harbor. 
As shown in the plot, Chrysene (569,679.08 ng/g) has the highest concentration in the harbor. This is followed by Pyrene (346679.56 ng/g), Fluoranthene (319594.32 ng/g), and Phenanthrene (221,665.68 ng/g). 

```{r sum-organics-boston-inner-harbor, message=FALSE, echo=FALSE}
PAHs %>%
  group_by(GEN_LOC_NM, pah_detected) %>%
  summarise(amount_detected = sum(amount_detected, na.rm = TRUE)) %>%
  filter(GEN_LOC_NM == "Boston Inner Harbor") %>%
  arrange(desc(amount_detected))
``` 
``` {r sum-organics-boston-inner-harbor-graph}
PAHs %>%
  group_by(GEN_LOC_NM, pah_detected) %>%
  summarise(amount_detected = sum(amount_detected, na.rm = TRUE)) %>%
  filter(GEN_LOC_NM == "Boston Inner Harbor") %>%
 ggplot(aes(x = fct_rev(fct_reorder(pah_detected, amount_detected)), y = amount_detected, fill = pah_detected))+
  geom_col()+
  coord_flip()+
  theme_minimal() +
  theme(legend.position = "none")+
  labs(title = "PAH Concentrations in the Gulf of Maine",
       subtitle = "in Boston Inner Harbor",
       x = "PAH Detected",
       y = "Amount Detected (ng/g)")
```

### Northwest Boston Harbor Concentrations
This plot visualizes the specific PAHs present in Northwest Boston Harbor.  
As shown in the plot, Pyrene (315,904.49 ng/g) has the highest concentration in the harbor. This is followed by Fluoranthene (312,942.98 ng/g), Phenanthrene (219,061.23 ng/g), and Chrysene (200,784.37 ng/g).

```{r type_of_organic-in-northwest-boston-harbor, echo=FALSE, message=FALSE}
PAHs %>%
  group_by(GEN_LOC_NM, pah_detected) %>%
  summarise(amount_detected = sum(amount_detected, na.rm = TRUE)) %>%
  filter(GEN_LOC_NM == "Northwest Boston Harbor") %>%
  arrange(desc(amount_detected))
```
``` {r type_of_organic-in-northwest-boston-harbor-graph, warning = FALSE}
PAHs %>%
  drop_na(amount_detected) %>%
  filter(GEN_LOC_NM %in% c("Northwest Boston Harbor")) %>%
  ggplot(aes(x = fct_rev(fct_reorder(pah_detected, amount_detected)), y = amount_detected, fill = pah_detected))+
  geom_col()+
  coord_flip()+
  theme(legend.position = "none")+
  labs(title = "PAH Concentrations in Northwest Boston Harbor",
       x = "PAH Detected",
       y = "Amount Detected (ng/g)")
```

## Central Boston Harbor Concentrations
This plot visualizes the specific PAHs present from Central Boston Harbor.  
As shown in the plot, Pyrene (71,710.28 ng/g) has the highest concentration in the harbor. This is followed by Fluoranthene (64,217.22 ng/g), and Phenanthrene (43304.62 ng/g). 
``` {r sum-organics-boston-central-harbor, echo=FALSE, message=FALSE}
PAHs %>%
  group_by(GEN_LOC_NM, pah_detected) %>%
  summarise(amount_detected = sum(amount_detected, na.rm = TRUE)) %>%
  filter(GEN_LOC_NM == "Central Boston Harbor") %>%
  arrange(desc(amount_detected))
```
```{r sum-organics-boston-central-harbor-graph, warning=FALSE}
PAHs %>%
  drop_na(amount_detected) %>%
  filter(GEN_LOC_NM %in% c("Central Boston Harbor")) %>%
 ggplot(aes(x = fct_rev(fct_reorder(pah_detected, amount_detected)), y = amount_detected, fill = pah_detected))+
  geom_col()+
  coord_flip()+
  theme(legend.position = "none")+
  labs(title = "PAH Concentrations in Central Boston Harbor",
       x = "PAH Detected",
       y = "Amount Detected (ng/g)")
```

## Cape Ann to Cape Elizabeth Concentrations
This plot visualizes the specific PAHs present from Cape Ann to Cape Elizabeth.  
As shown in the plot, Fluoranthene (77,701.85 ng/g) has the highest concentration in the harbor. This is followed by Pyrene	(73,067.06 ng/g), Phenanthrene (46,915.83 ng/g), and Chrysene (35,656.56 ng/g). 

```{r type_of_organic-cape-ann-to-elizabeth, echo=FALSE, message=FALSE}
PAHs %>%
  group_by(GEN_LOC_NM, pah_detected) %>%
  summarise(amount_detected = sum(amount_detected, na.rm = TRUE)) %>%
  filter(GEN_LOC_NM == "Cape Ann to Cape Elizabeth") %>%
  arrange(desc(amount_detected))
```
```{r type_of_organic-cape-ann-to-elizabeth-graph, warning=FALSE}
PAHs %>%
  drop_na(amount_detected) %>%
  filter(GEN_LOC_NM %in% c("Cape Ann to Cape Elizabeth")) %>%
 ggplot(aes(x = fct_rev(fct_reorder(pah_detected, amount_detected)), y = amount_detected, fill = pah_detected))+
  geom_col()+
  coord_flip()+
  theme_minimal() +
  scale_fill_viridis_d() +
  theme(legend.position = "none")+
  labs(title = "PAH Concentrations from Cape Ann to Cape Elizabeth",
       x = "PAH Detected",
       y = "Amount Detected (ng/g)")
```

## Cape Elizabeth to Rockland Concentrations
This plot visualizes the specific PAHs present from Cape Elizabeth to Rockland.  
As shown in the plot, Fluoranthene (116,706.129 ng/g) has the highest concentration in the harbor. This is followed by Chrysene (112,928.05 ng/g), Pyrene (110,687.928 ng/g), and Phenanthrene (40,140.04 ng/g). 

```{r type_of_organic-in-cape-elizabeth-to-rockland, echo=FALSE, message=FALSE}
PAHs %>%
  group_by(GEN_LOC_NM, pah_detected) %>%
  summarise(amount_detected = sum(amount_detected, na.rm = TRUE)) %>%
  filter(GEN_LOC_NM == "Cape Elizabeth to Rockland") %>%
  arrange(desc(amount_detected))
```
```{r type_of_organic-in-cape-elizabeth-to-rockland-graph, warning=FALSE}
PAHs %>%
  drop_na(amount_detected) %>%
  filter(GEN_LOC_NM %in% c("Cape Elizabeth to Rockland")) %>%
 ggplot(aes(x = fct_rev(fct_reorder(pah_detected, amount_detected)), y = amount_detected, fill = pah_detected))+
  geom_col()+
  coord_flip()+
  theme_minimal() +
  scale_fill_viridis_d() +
  theme(legend.position = "none")+
  labs(title = "PAH Concentrations from Cape Elizabeth to Rockland",
       x = "PAH Detected",
       y = "Amount Detected (ng/g)")

PAHs %>%
  drop_na(amount_detected) %>%
  filter(GEN_LOC_NM %in% c("Cape Elizabeth to Rockland")) %>%
ggplot(aes(x="", y = amount_detected, fill = pah_detected)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0)
```


## Statistical tests
### Difference between general locations

Does total PAH concentration differ significantly between general locations?

```{r}
hist(PAHs$amount_detected)
```

```{r}
kw1=kruskal.test(PAHs$amount_detected ~ PAHs$GEN_LOC_NM)
kw1
```
This shows that it is significant. We will now run a Dunn post-hoc test to identify any significance between locations.
```{r}
#install.packages("dunn.test")
#library(dunn.test)

#dunn=dunn.test(PAHs$amount_detected,Organics$GEN_LOC_NM,method="bh")

```
Interpretation: Most general locations have significantly different mean total PAH concentrations.

## Map Plots
### Static maps
```{r load-GOM_states-shp}
GOM_states <- st_read("/cloud/project/extra/GOM_DD.shp")
```

```{r load-bathy-shp}
Bathy <- st_read("/cloud/project/extra/BATHYMGM_ARC.shp")
```
Bathymetry data is in projected coordinate system (NAD83), other data is geodetic (NAD83).  
Wasn't able to add this layer to map plot -- need to change projection.

```{r bathy-to-NAD83-format}
st_crs(Bathy)
Bathy <- st_transform(Bathy, "+init=epsg:4269")
```
Now the Bathy shapefile is in geodetic NAD83 format.

```{r unique-bathy}
unique(Bathy$CONTOUR)
```

```{r plot-bathy}
Bathy_low_res <- Bathy%>%
 filter(CONTOUR %in% c("-100","-500","-1000","-2000","-3000","-4000"))
ggplot(Bathy_low_res) +
  geom_sf(aes())
```
```{r creating-Bathy_hi_res-pah}
Bathy_hi_res <- Bathy%>%
 filter(CONTOUR %in% c("-40","-80","-120","160","200","240"))
ggplot(Bathy_hi_res) +
  geom_sf(aes())
```

```{r drop-na-pah}
PAH_no_na <- PAHs %>%
  drop_na(amount_detected)

PAH_no_na_specific_pah <- PAH_no_na %>%
  filter(pah_detected %in% c("BENZNE", "BENZNE", "NAPHTHLN", "BIPHENYL", "ACNPHTHN", "ACNPHTYL", "FLUORENE", "PHNANTHR", "ANTHRACN", "PYRENE", "FLORNTHN", "CHRYSENE", "PERYLENE"))

```


```{r pah-gom-map-plot, warning = FALSE}
ggplot(GOM_states)+
  geom_sf(aes())+
  geom_sf(data = Bathy_low_res, color = "gray80", width = 1)+
  geom_point(data = PAH_no_na, (aes(x = LONGITUDE, y = LATITUDE, size = amount_detected, alpha = 0.5)))+
  xlim(-72,-65)+
  ylim(40,45)+
  theme_bw()+
  labs(title = "Distribution and concentration of PAHs",
       subtitle ="Gulf of Maine sediments",
       x = "Longitude",
       y = "Latitude")+
  guides(size = guide_legend(title = "PAH ng/g"))+
  guides(alpha = FALSE)+
  ggspatial::annotation_scale(
    location = "bl",
    bar_cols = c("grey60", "white"),
    text_family = "ArcherPro Book"
  ) +
  ggspatial::annotation_north_arrow(
    location = "tr", which_north = "true",
    pad_x = unit(0, "in"), pad_y = unit(0.2, "in"),
    style = ggspatial::north_arrow_nautical(
      fill = c("grey40", "white"),
      line_col = "grey20",
      text_family = "ArcherPro Book"))
```
Above is the Distribution and concentration of PAHs in the Gulf of Maine sediments. 

```{r MDI-area-map-pah}
ggplot(GOM_states)+
  geom_sf(aes())+
  geom_sf(data = Bathy_hi_res, color = "gray80", width = 1)+
  geom_point(data=PAH_no_na, (aes(x = LONGITUDE, y = LATITUDE, size = amount_detected, alpha = 0.5)))+
  xlim(-69.2,-68)+
  ylim(44,44.5)+
  theme_bw()+
  labs(title = "Distribution and concentration of PAHs",
       subtitle ="MDI and Penobscot Bay sediments",
       x = "Longitude",
       y = "Latitude")+
  guides(size = guide_legend(title = "PAH ng/g"))+
  guides(alpha = FALSE)+
  ggspatial::annotation_scale(
    location = "bl",
    bar_cols = c("grey60", "white"),
    text_family = "ArcherPro Book"
  ) +
  ggspatial::annotation_north_arrow(
    location = "tr", which_north = "true",
    pad_x = unit(0, "in"), pad_y = unit(0.2, "in"),
    style = ggspatial::north_arrow_nautical(
      fill = c("grey40", "white"),
      line_col = "grey20",
      text_family = "ArcherPro Book"))
```
Above is the Distribution and concentration of PAHs in MDI and Penobscot Bay sediments. 

```{r boston-and-ma-area-map-pah}
ggplot(GOM_states)+
  geom_sf(aes())+
  geom_sf(data = Bathy_hi_res, color = "gray80", width = 1)+
  geom_point(data=PAH_no_na, (aes(x = LONGITUDE, y = LATITUDE, size = amount_detected, alpha = 0.5)))+
  xlim(-71.2,-69.5)+
  ylim(41.8,43)+
  theme_bw()+
  labs(title = "Distribution and concentration of PAHs",
       subtitle ="Boston, MA and Cape Cod Bay sediments",
       x = "Longitude",
       y = "Latitude")+
  guides(size = guide_legend(title = "PAH ng/g"))+
  guides(alpha = FALSE)+
  ggspatial::annotation_scale(
    location = "bl",
    bar_cols = c("grey60", "white"),
    text_family = "ArcherPro Book"
  ) +
  ggspatial::annotation_north_arrow(
    location = "tr", which_north = "true",
    pad_x = unit(0, "in"), pad_y = unit(0.1, "in"),
    style = ggspatial::north_arrow_nautical(
      fill = c("grey40", "white"),
      line_col = "grey20",
      text_family = "ArcherPro Book"))
```
Above is the Distribution and concentration of PAHs in Boston, MA and Cape Cod Bay sediments. 

```{r boston-and-ma-area-map-pah-colored}
ggplot(GOM_states)+
  geom_sf(aes())+
  geom_sf(data = Bathy_hi_res, color = "gray80", width = 1)+
  geom_point(data=PAH_no_na_specific_pah, (aes(x = LONGITUDE, y = LATITUDE, size = amount_detected, color = pah_detected, alpha = 0.5)))+
  xlim(-71.2,-69.5)+
  ylim(41.8,43)+
  theme_bw()+
  labs(title = "Distribution and concentration of PAHs",
       subtitle ="Boston, MA and Cape Cod Bay sediments",
       x = "Longitude",
       y = "Latitude")+
  guides(size = guide_legend(title = "PAH ng/g"))+
  guides(alpha = FALSE)+
  ggspatial::annotation_scale(
    location = "bl",
    bar_cols = c("grey60", "white"),
    text_family = "ArcherPro Book"
  ) +
  ggspatial::annotation_north_arrow(
    location = "tr", which_north = "true",
    pad_x = unit(0, "in"), pad_y = unit(0.1, "in"),
    style = ggspatial::north_arrow_nautical(
      fill = c("grey40", "white"),
      line_col = "grey20",
      text_family = "ArcherPro Book"))
```
Above is the same graph, but now we are coloring the points by the specific pah that was detected. 

### Interactive map

```{r create-html-labels-pah}
#labels <- sprintf("<strong>%s</strong><br/>%g ng/g", 
#                 PAH_no_na$SPECFC_LOC, PAH_no_na$amount_detected) %>%
# lapply(htmltools::HTML)
#
#head(labels, 1)
```

```{r leaflet-map-pcb}
#leaflet(data = PAH_no_na) %>%
  #addProviderTiles(providers$Esri.WorldTopoMap) %>%
  #setView(lng = -68.5, 
    #      lat = 43.5, 
     #     zoom = 6) %>%
  #addCircleMarkers(lng = ~LONGITUDE, lat = ~LATITUDE, popup = c(~SPECFC_LOC, ~amount_detected), label = labels)
```

Note that the two above code chunks are commented out as .rmd will not knit to github document with html functions.  Uncomment to run and change output type to `html_document` to knit.


