---
title: 'GOM Contaminated Sediments analysis: PCBs'
author: "Joshua Harkness and Autumn Pauly"
date: "2023-10-28"
output: github_document
---

```{r load-packages}
library(tidyverse)
library(sf)
library(leaflet)
library(RColorBrewer)
#install.packages("dplyr")
library(dplyr)
```

```{r load-data}
PCBs <- read.csv(paste0("/cloud/project/analysis/PCBs_loc.csv"), header = T)
Organics <- read.csv(paste0("/cloud/project/analysis/Organics_loc.csv"), header = T)
```

```{r glimpse-data, warning=FALSE}
glimpse(PCBs)
```

```{r combine-variables}
#PCBs %>%
  #drop_na(c(PCB_52_NGG:PCB209_NGG))
  #mutate(PCBs = sum(c(PCB_52_NGGPCB209_NGG)), na.rm = T)
  #unite("PCBs", PCB_52_NGG:PCB209_NGG, remove = TRUE)

##This is what I (Autumn) tried to do, but it still wouldn't recognize the columns. 
#PCBs %>%
  #drop_na(c(PCB_52_NGG:PCB209_NGG))
  #mutate(PCBs = PCB_52_NGG + PCB101_NGG + PCB118_NGG + PCB128_NGG + PCB138_NGG + PCB153_NGG + PCB180_NGG + PCB206_NGG + PCB209_NGG), na.rm = T)
  #unite("PCBs", PCB_52_NGG:PCB209_NGG, remove = TRUE)
```

```{r joining-columns, warning=FALSE}

#PCBs <- PCBs %>%
#  mutate(sum_PCBs = PCB101_NGG + PCB118_NGG + PCB128_NGG + PCB138_NGG + PCB153_NGG + PCB180_NGG + PCB206_NGG + PCB209_NGG) %>% 
 # mutate(sumBHCs = BHC_A_C + BHC_B_C + BHC_D_C)

#PCBs <- PCBs %>% 
#  select(-BHC_A_C, -BHC_B_C, -BHC_D_C, -PCB101_NGG,  -PCB118_NGG, -PCB128_NGG, -PCB138_NGG, -PCB153_NGG, #-PCB180_NGG, -PCB206_NGG, -PCB209_NGG)

#glimpse(PCBs)

```

```{r pcbs-sum-rowwise}
#PCBs <- PCBs %>%
#  rowwise() %>%
#mutate(PCBs_sum_total = sum(across(c(PCB_52_NGG, PCB101_NGG, PCB118_NGG, PCB128_NGG, PCB138_NGG, #PCB153_NGG, PCB180_NGG, PCB206_NGG, PCB209_NGG)), na.rm = T))

#PCBs %>%
#  group_by(GEN_LOC_NM) %>%
#  select(STATE_NAME, GEN_LOC_NM, SPECFC_LOC, PCBs_sum_total) %>%
#  arrange(desc(PCBs_sum_total))
```

 
```{r calculate-sum-stats-organics}
Summary_Organics <- Organics %>%
  group_by(GEN_LOC_NM) %>%
  drop_na(PCB_T_UGG) %>%
  summarise(mean_PCB_T = mean(PCB_T_UGG),
    sd_PCB_T = sd(PCB_T_UGG),
    n_PCB_T = n(),
    SE_PCB_T = sd(PCB_T_UGG) / sqrt(n()))

Summary_Organics
```

```{r pcb-gen-loc}
Summary_Organics %>%
  ggplot(aes(x = fct_rev(fct_reorder(GEN_LOC_NM, mean_PCB_T)), y = mean_PCB_T, fill = GEN_LOC_NM))+
  geom_col(col = "black")+
  geom_errorbar(aes(ymin = mean_PCB_T - SE_PCB_T, ymax = mean_PCB_T + SE_PCB_T), width = 0.2)+
  coord_flip()+
  scale_fill_viridis_d()+
  theme_bw()+
  theme(legend.position = "none")+
  labs(title = "PCB Concentrations in Gulf of Maine Sediments",
       x = "General location",
       y = "Mean total PCB concentration ug/g",
       caption = "Error bars = 1 standard error")
```
 
 
```{r pcb-boston-harbor1}
Summary_Organics %>%
  filter(GEN_LOC_NM %in% c("BOSTON INNER HARBOR", "CENTRAL BOSTON HARBOR", "SOUTHEAST BOSTON HARBOR", "NORTHWEST BOSTON HARBOR")) %>%
  ggplot(aes(x = fct_rev(fct_reorder(GEN_LOC_NM, mean_PCB_T)), y = mean_PCB_T, fill = GEN_LOC_NM))+
  geom_col(col = "black")+
  geom_errorbar(aes(ymin = mean_PCB_T - SE_PCB_T, ymax = mean_PCB_T + SE_PCB_T), width = 0.2)+
  scale_fill_brewer(type = "qual", palette = 4, direction = 1, aesthetics = "fill")+
  theme_bw()+
  coord_flip()+
  theme(legend.position = "none")+
  labs(title = "PCB Concentrations in Boston Harbor Sediments",
       x = "General location",
       y = "Mean total PCB concentration ug/g",
       caption = "Error bars = 1 standard error")
```
 
```{r pcb-boston-harbor2}
Summary_Organics %>%
  filter(GEN_LOC_NM %in% c("CENTRAL BOSTON HARBOR", "NORTHWEST BOSTON HARBOR")) %>%
  ggplot(aes(x = fct_reorder(GEN_LOC_NM, mean_PCB_T), y = mean_PCB_T, fill = GEN_LOC_NM))+
  geom_col(col = "black")+
  geom_errorbar(aes(ymin = mean_PCB_T - SE_PCB_T, ymax = mean_PCB_T + SE_PCB_T), width = 0.2)+
  scale_fill_brewer(type = "qual", palette = 4, direction = 1, aesthetics = "fill")+
  theme_bw()+
  theme(legend.position = "none")+
  labs(title = "PCB Concentrations in Boston Harbor Sediments",
       subtitle = "Central and Northwest Boston Harbor",
       x = "General location",
       y = "Mean total PCB concentration ug/g",
       caption = "Error bars = 1 standard error")
```

### Specific Locations


```{r}
Sum_Org_site <- Organics %>%
  group_by(site) %>%
  drop_na(PCB_T_UGG) %>%
  summarise(mean_PCB_T = mean(PCB_T_UGG),
    sd_PCB_T = sd(PCB_T_UGG),
    n_PCB_T = n(),
    SE_PCB_T = sd(PCB_T_UGG) / sqrt(n()))

Sum_Org_site %>%
  distinct(site)
```



```{r}
Sum_Org_site %>%
  filter(site %in% c("In Piscataqua River NE of Outer Cutts Cove", "Piscataqua River N of Pierce Is.","York River", "Kennebunk River", "Portland Fore River", "Kennebec River to Bath", "Penobscot River To Bangor")) %>%
  ggplot(aes(x = fct_rev(fct_reorder(site, mean_PCB_T)), y = sapply(mean_PCB_T, FUN=function(x) ifelse(x==0.000000e0, -0.01,x) ), fill = site))+
  geom_bar(stat="identity", col = "black")+
  scale_x_discrete(drop=FALSE)+
  geom_errorbar(aes(ymin = mean_PCB_T - SE_PCB_T, ymax = mean_PCB_T + SE_PCB_T), width = 0.2)+
  scale_fill_brewer(type = "qual", palette = 4, direction = 1, aesthetics = "fill")+
  theme_bw()+
  coord_flip()+
  theme(legend.position = "none")+
  labs(title = "PCB Concentrations in Maine Rivers",
       subtitle = "Union River excluded",
       caption = "Error bars = 1 standard error",
       x = "Location",
       y = "Mean total PCB concentration ug/g")
```

Need to figure out how to keep categories with zero values.  For example, samples from the Penobscot River yielded no PCBs and are entered as zero values but will not display in ggplot.  Tried "scale_x_discrete(drop = FALSE)" with no effect.

Also would like to combine the values for the two Piscataqua River sites into one category, so that there is only one bar in these plots.

```{r}
Sum_Org_site %>%
  filter(site %in% c("In Piscataqua River NE of Outer Cutts Cove", "Piscataqua River N of Pierce Is.","York River", "Kennebunk River", "Portland Fore River", "Kennebec River to Bath", "Penobscot River To Bangor", "Union River")) %>%
  ggplot(aes(x = fct_rev(fct_reorder(site, mean_PCB_T)), y = sapply(mean_PCB_T, FUN=function(x) ifelse(x==0.000000e0, -2,x)), fill = site))+
  geom_col(col = "black")+
  geom_errorbar(aes(ymin = mean_PCB_T - SE_PCB_T, ymax = mean_PCB_T + SE_PCB_T), width = 0.2)+
  scale_fill_manual(values = c("gray50", "skyblue", "skyblue", "gray50", "gray50", "gray50", "chocolate1", "gray50"))+
  theme_bw()+
  coord_flip()+
  theme(legend.position = "none")+
  labs(title = "PCB Concentrations in Maine Rivers",
       subtitle = "With Union River, Ellsworth",
       caption = "Error bars = 1 standard error",
       x = "Location",
       y = "Mean total PCB concentration ug/g")
```

There are no errorbars for Union River as n = 1; single sample from this site so only a mean value.

```{r}
Sum_Org_site %>%
  filter(site %in% c("BASS RIVER", "ESSEX RIVER", "MYSTIC RIVER", "FORE RIVER", "CHELSEA RIVER", "Chelsea River", "Mill Creek", "Neponset River Bridge", "South River", "Weymouth Fore & Town River", "North & Danvers River")) %>%
  ggplot(aes(x = fct_rev(fct_reorder(site, mean_PCB_T)), y = mean_PCB_T, fill = site))+
  geom_bar(stat="identity", col = "black")+
  scale_x_discrete(drop=FALSE)+
  geom_errorbar(aes(ymin = mean_PCB_T - SE_PCB_T, ymax = mean_PCB_T + SE_PCB_T), width = 0.2)+
  scale_fill_brewer(type = "qual", palette = 4, direction = 1, aesthetics = "fill")+
  theme_bw()+
  coord_flip()+
  theme(legend.position = "none")+
  labs(title = "PCB Concentrations in Massachusetts Rivers",
       caption = "Error bars = 1 standard error",
       x = "Location",
       y = "Mean total PCB concentration ug/g")
```


## Statistical tests
### Difference between general locations

Does total PCB concentration differ significantly between general locations?

```{r}
hist(Organics$PCB_T_UGG)
```
The distribution of total PCB concentrations is highly skewed, with high influence outliers.  It fails to meet parametric assumptions, so it should be analyzed with non-parametric methods.  We could do bootstrapping here, but a Kruskal Wallis test is appropriate and simpler.

```{r}
kw1=kruskal.test(Organics$PCB_T_UGG ~ Organics$GEN_LOC_NM)
kw1
```

Significant -- now a Dunn post-hoc test to identify significance between locations.

```{r}
#install.packages("dunn.test")
library(dunn.test)

dunn=dunn.test(Organics$PCB_T_UGG,Organics$GEN_LOC_NM,method="bh")

```
Interpretation: Most general locations have significantly different mean total PCB concentrations...



## Map Plots
### Static maps
```{r load-GOM_states-shp}
GOM_states <- st_read("/cloud/project/extra/GOM_DD.shp")
```
```{r load-bathy-shp}
Bathy <- st_read("/cloud/project/extra/BATHYMGM_ARC.shp")
```

Bathymetry data is in projected coordinate system (NAD83), other data is geodetic (NAD83).  Wasn't able to add this layer to map plot -- need to change projection.


```{r}
st_crs(Bathy)
Bathy <- st_transform(Bathy, "+init=epsg:4269")
```


Now the Bathy shapefile is in geodetic NAD83 format.

```{r unique-bathy}
unique(Bathy$CONTOUR)
```



```{r plot-bathy}
Bathy_low_res <- Bathy%>%
 filter(CONTOUR %in% c("-100","-500","-1000","-2000","-3000","-4000"))
ggplot(Bathy_low_res) +
  geom_sf(aes())
```
```{r}
Bathy_hi_res <- Bathy%>%
 filter(CONTOUR %in% c("-40","-80","-120","160","200","240"))
ggplot(Bathy_hi_res) +
  geom_sf(aes())
```


```{r drop-na-PCB_T}
Org_no_na <- Organics %>%
  drop_na(PCB_T_UGG)
```
```{r pcb-gom-map-plot, warning = FALSE}
ggplot(GOM_states)+
  geom_sf(aes())+
  geom_sf(data = Bathy_low_res, color = "gray80", width = 1)+
  geom_point(data = Org_no_na, (aes(x = LONGITUDE, y = LATITUDE, size = PCB_T_UGG, alpha = 0.5)))+
  xlim(-72,-65)+
  ylim(40,45)+
  theme_bw()+
  labs(title = "Distribution and concentration of PCBs",
       subtitle ="Gulf of Maine sediments",
       x = "Longitude",
       y = "Latitude")+
  guides(size = guide_legend(title = "PCB ug/g"))+
  guides(alpha = FALSE)+
  ggspatial::annotation_scale(
    location = "bl",
    bar_cols = c("grey60", "white"),
    text_family = "ArcherPro Book"
  ) +
  ggspatial::annotation_north_arrow(
    location = "tr", which_north = "true",
    pad_x = unit(0, "in"), pad_y = unit(0.2, "in"),
    style = ggspatial::north_arrow_nautical(
      fill = c("grey40", "white"),
      line_col = "grey20",
      text_family = "ArcherPro Book"))
```

```{r MDI-area-map}
ggplot(GOM_states)+
  geom_sf(aes())+
  geom_sf(data = Bathy_hi_res, color = "gray80", width = 1)+
  geom_point(data=Org_no_na, (aes(x = LONGITUDE, y = LATITUDE, size = PCB_T_UGG, alpha = 0.5)))+
  xlim(-69.2,-68)+
  ylim(44,44.5)+
  theme_bw()+
  labs(title = "Distribution and concentration of PCBs",
       subtitle ="MDI and Penobscot Bay sediments",
       x = "Longitude",
       y = "Latitude")+
  guides(size = guide_legend(title = "PCB ug/g"))+
  guides(alpha = FALSE)+
  ggspatial::annotation_scale(
    location = "bl",
    bar_cols = c("grey60", "white"),
    text_family = "ArcherPro Book"
  ) +
  ggspatial::annotation_north_arrow(
    location = "tr", which_north = "true",
    pad_x = unit(0, "in"), pad_y = unit(0.2, "in"),
    style = ggspatial::north_arrow_nautical(
      fill = c("grey40", "white"),
      line_col = "grey20",
      text_family = "ArcherPro Book"))
```

```{r}
ggplot(GOM_states)+
  geom_sf(aes())+
  geom_sf(data = Bathy_hi_res, color = "gray80", width = 1)+
  geom_point(data=Org_no_na, (aes(x = LONGITUDE, y = LATITUDE, size = PCB_T_UGG, alpha = 0.5)))+
  xlim(-71.2,-69.5)+
  ylim(41.8,43)+
  theme_bw()+
  labs(title = "Distribution and concentration of PCBs",
       subtitle ="Boston, MA and Cape Cod Bay sediments",
       x = "Longitude",
       y = "Latitude")+
  guides(size = guide_legend(title = "PCB ug/g"))+
  guides(alpha = FALSE)+
  ggspatial::annotation_scale(
    location = "bl",
    bar_cols = c("grey60", "white"),
    text_family = "ArcherPro Book"
  ) +
  ggspatial::annotation_north_arrow(
    location = "tr", which_north = "true",
    pad_x = unit(0, "in"), pad_y = unit(0.1, "in"),
    style = ggspatial::north_arrow_nautical(
      fill = c("grey40", "white"),
      line_col = "grey20",
      text_family = "ArcherPro Book"))
```


### Interactive map

```{r create-html-labels}
#labels <- sprintf("<strong>%s</strong><br/>%g ug/g", 
#                  Org_no_na$SPECFC_LOC, Org_no_na$PCB_T_UGG) %>%
#  lapply(htmltools::HTML)

#head(labels, 1)
```

```{r leaflet-map-pcb}
#leaflet(data = Org_no_na) %>%
#  addProviderTiles(providers$Esri.WorldTopoMap) %>%
#  setView(lng = -68.5, 
#          lat = 43.5, 
#          zoom = 6) %>%
#  addCircleMarkers(lng = ~LONGITUDE, lat = ~LATITUDE, popup = c(~SPECFC_LOC, ~PCB_T_UGG), label = labels)
```

Note that the two above code chunks are commented out as .rmd will not knit to github document with html functions.  Uncomment to run and change output type to `html_document` to knit.
